---
title: "Chi Square Test"
output: html_notebook
---

### Introduction
In order to establish that 2 categorical variables are dependent, the chi-squared statistic should be above a certain cutoff. This cutoff increases as the number of classes within the variable increases.

Alternatively, we can just perform a chi-squared test and check the `p-values`.

Like all statistical tests, chi-squared test assumes a null hypothesis and an alternate hypothesis. The general practice is, if the p-value that comes out in the result is less than a pre-determined significance level, which is `0.05` usually, then we reject the null hypothesis.

**H0**: The The two variables are independent

**H1**: The two variables are related.

The null hypothesis of the chi-squared test is that the two variables are independent and the alternate hypothesis is that they are related.

### Who is smoking more
We are going to use `survey` dataset in `MASS` package to implement an example of *Chi Squre test*. We will see the relation between the `Smoke` feature observations of the students smoking habit and the gender in `sex` column to see if gender affects this bad habit.

First, let us look at the data
```{r}
library(MASS)
str(survey)
```

Now let us create the *Contingency table*

```{r}
gender_smokers <- table(survey$Sex, survey$Smoke)
```


Now we can perform the test on the resultant contingency table.

```{r}
chi <- chisq.test(gender_smokers)
print(chi)
```

We have a chi-squared value of `r chi$p.value`. Since we get a p-Value greater than the significance level of `0.05`, we accept the null hypothesis and conclude that the two variables are in fact independent. 